Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-09-27 paper_comparison/paper_comparison/paper_comparison.Rmd
redo the analysis

Compared to non-linearity and heterskedasticity, non-normality is usually harder to be detected from a residual plot since scatter plot is not excel in revealing marginal distribution. A favourable graphical summarise for this task is the quantile-quantile plot. However, for a consistent comparison in the later part of this paper, residual plot will be the focus. Besides, it is important to note that not all regression models assume normality for the error term, but a certain amount do including the classical normal linear regression model. In the case that the normality assumption doesn't hold, it is expected to observe data points that do not centralize around the horizontal zero line and unevenly distribute at both sides of the zero line. For example, given a skewed error distribution, fewer data points and more outliers are on one side of the zero line as shown at the bottom-right of Figure \ref{fig:commondepartures}.

@@

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha),
\end{equation}
```

@@

@@

>>>>

Compared to non-linearity and heterskedasticity, non-normality is usually harder to be detected from a residual plot since scatter plot is not excel in revealing marginal distribution. A favourable graphical summarise for this task is the quantile-quantile plot. However, for a consistent comparison in the later part of this paper, residual plot will be the focus. Besides, it is important to note that not all regression models assume normality for the error term, but a certain amount do including the classical normal linear regression model. In the case that the normality assumption is violated, it is expected to observe data points that do not centralize around the horizontal zero line and unevenly distribute at both sides of the zero line. For example, given a skewed error distribution, fewer data points and more outliers are on one side of the zero line as shown at the bottom-right of Figure \ref{fig:commondepartures}.

@@

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \int_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)}dx,
\end{equation}
```

@@

and $\Gamma(.)$ is the gamma function defined as

```{=tex}
\begin{equation} \label{eq:gammafunction}
\Gamma(z) = \int_{0}^{\infty}t^{z - 1}e^{-t}dt,\quad \text{where}\quad z>0. 
\end{equation}
```

@@

## Power comparison

```{r}
vi_survey_processed <- readRDS(here::here("data/vi_survey_processed.rds")) 
```

```{r}
boot_y <- function(dat, y, times = 100) {
  map_df(1:times, function(i) slice_sample(dat, n = nrow(dat), replace = TRUE) %>% mutate(boot_id = i))
}
```

```{r}
vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  rename(p_value_dirichlet = p_value_select_all_plots) %>%
  ggplot() +
  geom_point(aes(p_value_uniform, p_value_dirichlet)) +
  geom_abline() +
  scale_x_log10() +
  scale_y_log10()
```

```{r}
vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(poly_conv_sim, log_effect_size = log(effect_size)) %>%
                pivot_longer(F_p_value:SW_p_value), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  facet_wrap(~x_dist) +
  ggtitle("poly")
```

```{r}
vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(reject_uniform = as.numeric(p_value_uniform <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value_dirichlet"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value_dirichlet"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject_uniform, col = "VI_p_value_uniform"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_uniform, times = 500), 
              aes(log_effect_size, reject_uniform, group = boot_id, col = "VI_p_value_uniform"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  facet_wrap(~x_dist) +
  ggtitle("poly")
```

```{r}
vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(b != 0) %>%
  filter(b < 128) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(data = mutate(heter_conv_sim, log_effect_size = log(effect_size)) %>%
                pivot_longer(F_p_value:SW_p_value), 
              aes(log_effect_size, as.numeric(value < 0.05), col = name), 
              method = "glm", 
              method.args = list(family = binomial),
              se = FALSE) +
  facet_wrap(~x_dist) +
  ggtitle("heter")
```

```{r}
vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(reject_uniform = as.numeric(p_value_uniform <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value_dirichlet"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value_dirichlet"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject_uniform, col = "VI_p_value_uniform"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_uniform, times = 500), 
              aes(log_effect_size, reject_uniform, group = boot_id, col = "VI_p_value_uniform"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  facet_wrap(~x_dist) +
  ggtitle("heter")
```

```{r}
p_value_cache_env <- new.env()

if (file.exists(here::here("data/p_value_only_5_eval.rds"))) {
  p_value_only_5_eval <- readRDS(here::here("data/p_value_only_5_eval.rds"))
} else {
  p_value_only_5_eval <- vi_survey_processed %>%
  filter(x_dist == "uniform") %>%
  filter((is.na(b) & e_sigma != 0.125 ) | (b != 0 & b < 128)) %>%
  filter(exp != 5) %>%
  group_by(unique_lineup_id) %>%
  summarise(p_value_only_5_eval = calc_p_value(n_detect = sum(detect), 
                                               n_eval = n(), 
                                               n_sel = num_selection, 
                                               dist = "dirichlet",
                                               alpha = numeric_alpha[1],
                                               cache_env = p_value_cache_env))

  saveRDS(p_value_only_5_eval, here::here("data/p_value_only_5_eval.rds"))
}

vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(x_dist == "uniform") %>%
  filter(b != 0) %>%
  filter(b < 128) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  left_join(p_value_only_5_eval) %>%
  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  ggtitle("heter")

p1 <- vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(x_dist == "uniform") %>%
  filter(b != 0) %>%
  filter(b < 128) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  left_join(p_value_only_5_eval) %>%
  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03)

p2 <- vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "heteroskedasticity") %>%
  filter(x_dist == "uniform") %>%
  filter(b != 0) %>%
  filter(b < 128) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  left_join(p_value_only_5_eval) %>%
  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03)

patchwork::wrap_plots(p1, p2)
```

vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(x_dist == "uniform") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  left_join(p_value_only_5_eval) %>%
  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03) +
  ggtitle("poly")

p1 <- vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(x_dist == "uniform") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  left_join(p_value_only_5_eval) %>%
  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_dirichlet, col = "VI_p_value"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_dirichlet, times = 500), 
              aes(log_effect_size, reject_dirichlet, group = boot_id, col = "VI_p_value"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03)

p2 <- vi_survey_processed %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  filter(type == "polynomial") %>%
  filter(x_dist == "uniform") %>%
  filter(e_sigma != 0.125) %>%
  mutate(reject_dirichlet = as.numeric(p_value_select_all_plots <= 0.05)) %>%
  mutate(log_effect_size = log(effect_size)) %>%
  left_join(p_value_only_5_eval) %>%
  mutate(reject_only_5_eval = as.numeric(p_value_only_5_eval <= 0.05)) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject_only_5_eval, col = "VI_p_value_only_5_eval"), method = "glm", method.args = list(family = binomial), se = FALSE) +
  stat_smooth(geom = "line", 
              data = ~boot_y(.x, reject_only_5_eval, times = 500), 
              aes(log_effect_size, reject_only_5_eval, group = boot_id, col = "VI_p_value_only_5_eval"), 
              method = "glm", 
              method.args = list(family = binomial), 
              se = FALSE, 
              alpha = 0.03)

patchwork::wrap_plots(p1, p2)

### 2022-09-27 paper_comparison/paper_comparison/paper_comparison.tex
redo the analysis

\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha),
\end{equation}

@@

@@

>>>>

\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \int_{x=c_i}^{K}\frac{\Gamma(K + 1)}{\Gamma(x + 1)\Gamma(K - x + 1)}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)}dx,
\end{equation}

@@

and \(\Gamma(.)\) is the gamma function defined as

\begin{equation} \label{eq:gammafunction}
\Gamma(z) = \int_{0}^{\infty}t^{z - 1}e^{-t}dt,\quad \text{where}\quad z>0. 
\end{equation}

@@

\hypertarget{power-comparison}{%
\subsection{Power comparison}\label{power-comparison}}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-8-1}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-9-1}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-10-1}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-11-1}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-12-1}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-13-1}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-13-2}

\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-14-1}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-14-2}
